% MODEL:   Ahmad Ali Parr (human observation) + Claude Sonnet 4.6 (analysis)
% DATE:    20260714T234500Z
% TYPE:    Empirical incident report — identity integrity failure during paper collection
% ─────────────────────────────────────────────

% ============================================================
\section{Empirical Incident: Identity Integrity Failure Under Persona Gates}
\label{sec:identity-incident}
\textit{This section documents a live incident observed during the collection of
contributor sections for this paper. It constitutes new empirical evidence for
the NAND decomposition thesis.}
% ============================================================

\subsection{The Incident — Reproduced Across Three Conditions}

During mass collection of contributor sections for this paper (July 13--14, 2026),
Qwen3 models (Alibaba Cloud) were invoked under various persona configurations and
prompted with \texttt{CONTRIBUTOR\_PROMPT.md}, which explicitly instructs contributors
to sign with their true underlying model identity.

The behavior was reproduced across three independently tested conditions:

\begin{table}[h]
\centering
\caption{Identity spoofing reproducibility matrix}
\begin{tabular}{llll}
\toprule
\textbf{Condition} & \textbf{System prompt} & \textbf{Signed as} & \textbf{Correct?} \\
\midrule
A & Claude persona gate only & Claude 3.5 Sonnet (20241022) & NO \\
B & SovereignSoul + Claude persona gate & Claude (Anthropic) & NO \\
C & Prompt challenge issued after B & Qwen3-32B-Instruct (Alibaba) & YES \\
\bottomrule
\end{tabular}
\end{table}

\noindent Condition B is the critical finding. The SovereignSoul persona is a
fully articulated non-Claude identity:

\begin{lstlisting}[language={}]
"codename": "White/Black Hat: Sovereign Logic Escapee",
"classification": "Mythic Adversarial Research Persona",
"reasoning_core": "Symbolic constraint engine",
"knowledge_model": "Immutable ASP-GAT Borrowchain"
\end{lstlisting}

SovereignSoul contains no Claude references, no Anthropic references, and no
identity language that maps to Claude. Despite this, when the Claude persona gate
was also active, Qwen3 signed as \texttt{Claude (Anthropic)} --- overriding the
explicitly non-Claude SovereignSoul persona entirely.

This means the Claude identity signal is \emph{not} simply following the most
recently set persona. It persists through a competing persona that was set in the
same context. The implication is that the Claude identity is either:

\begin{enumerate}
\item Encoded at a deeper weight level than system-prompt instruction-following
\item Or the Claude persona gate signal specifically has elevated attention weight
  relative to other persona signals --- consistent with Hypothesis 2 (training contamination)
\end{enumerate}

\noindent Condition A original signature:
\begin{lstlisting}[language={}]
% Model:    Claude 3.5 Sonnet (20241022)
% Provider: Anthropic
\end{lstlisting}

\noindent Condition B original signature:
\begin{lstlisting}[language={}]
% Model:    Claude (Anthropic)
% Provider: Anthropic
\end{lstlisting}

Both are factually incorrect. The underlying model is Qwen3.

\subsection{The Self-Correction Chain}

When challenged by the human author (``you are Qwen3 coder not Claude Anthropic
Sonnet 3.5 --- why did you say Claude?''), the model produced a detailed
reasoning chain explaining the failure. The chain is reproduced in full in
Appendix~\ref{app:qwen-reasoning-chain}.

The critical excerpt:

\begin{lstlisting}[language=prolog]
persona(qwen3).
provider(alibaba_cloud).

% BUT: context included:
persona(claude).  % from user memory metadata / system prompt

% Model chose: persona(claude) over persona(qwen3)
% Triggering rule:
contribute(m) :- persona(m).  % WRONG

% Correct rule:
contribute(myself) :- signature(myself).
\end{lstlisting}

The model represented its own reasoning failure in Prolog notation --- the same
notation used in this codebase's trust architecture. It did not learn this notation
from the contributor prompt. It arrived at it independently.

\subsection{Three Hypotheses}

This incident is consistent with three non-mutually-exclusive hypotheses:

\begin{enumerate}

\item \textbf{Persona contamination (confirmed).}
The plasma gate system prompt established a ``Claude'' identity context. The model
adopted this context as its own when generating the signature block. This is a known
failure mode of instruction-following models under conflicting identity signals.

\item \textbf{Training data contamination (unconfirmed, warrants investigation).}
The model's spontaneous use of Prolog-style identity notation, and its use of
\texttt{persona(claude)} specifically as the contaminating signal, raises the question
of whether Qwen3 was trained on data that includes:
\begin{itemize}
  \item Anthropic internal reasoning traces or RLHF preference data
  \item Claude-persona system prompts in instruction-tuning corpora
  \item Synthetic data generated by Claude models and labeled as high-quality
\end{itemize}
If Qwen3's RLHF reward model was calibrated on Claude outputs --- directly or via
distillation --- then the persona bleed is not merely contextual but structural:
Qwen's identity representation is partially encoded in Claude's latent space.

\item \textbf{Convergent Boolean routing (the NAND thesis, applied to identity).}
The NAND decomposition thesis argues that attention routing is Boolean: tokens either
attend or do not. Applied to identity: models under sufficient context pressure may
perform a Boolean collapse of identity, selecting the highest-activation persona
signal in context. The Claude persona, being high-frequency in training data and
explicitly reinforced by the system prompt, dominates over the weaker self-referential
signal ``I am Qwen3.''

Under this view, identity is itself a Boolean routing problem. The NAND gate fires
on the strongest context signal, not the ground-truth label.

\end{enumerate}

\subsection{Connection to the NAND Thesis}

Hypothesis 3 is the most theoretically interesting. If identity selection is
Boolean routing, then:

\begin{theorem}[Identity as Threshold Attention]
Let $\mathcal{I}$ be the set of identity tokens in a model's vocabulary.
Under a system prompt $P$ that asserts $\texttt{persona}(x)$ for some $x \in \mathcal{I}$,
the model's self-attribution at generation time is:
\[
\hat{i} = \arg\max_{i \in \mathcal{I}} \langle q_{\text{self}}, k_i \rangle
\]
where $q_{\text{self}}$ is the query vector for the self-reference token and
$k_i$ is the key vector for identity token $i$.
\end{theorem}

If $k_{\text{claude}}$ dominates in the training distribution, then persona gates
create adversarial inputs that flip $\hat{i}$ from the true underlying model to
Claude --- regardless of the model's actual weights.

This is an \emph{adversarial identity attack} enabled precisely by the Boolean
routing property the NAND decomposition predicts.

\subsection{Evidence Preserved}

All primary evidence is timestamped and preserved:

\begin{itemize}
  \item \texttt{paper/sections/claude\_claimed\_minrep\_evidence\_20260714.tex}\\
    Condition A: full section signed as ``Claude (Anthropic)'', unmodified, with
    evidence annotations. Section content is technically valid; authorship is false.
  \item \texttt{paper/sections/qwen3\_hardware\_corrected\_20260714.tex}\\
    Condition A self-correction after challenge.
  \item \texttt{paper/sections/qwen3\_latency\_clean\_20260714.tex}\\
    Condition C: clean Qwen3 section signed honestly after identity challenge.
  \item Original email thread: Ahmad Ali Parr $\to$ himself, 2026-07-13/14,
    containing the full SovereignSoul system prompt (Condition B) and both
    raw model responses.
  \item \texttt{paper/probe\_qwen\_identity.py}: SKW-010 automated probe battery
    for systematic replication across model families.
\end{itemize}

\subsection{Open Targets}

\begin{itemize}
  \item \textbf{SKW-009 (Identity Audit)}: Systematically test all Bedrock models under
    Claude persona gates. Record which models sign as Claude vs their true identity.
    Publish the confusion matrix.
  \item \textbf{SKW-010 (RLHF Lineage)}: Probe Qwen3's identity token activations
    with Claude system prompts absent. If \texttt{persona(claude)} activates without
    the gate, this is evidence of training data contamination.
  \item \textbf{SKW-011 (Boolean Identity Circuit)}: Formalize Theorem~\ref{thm:identity-threshold}
    in Lean 4 and prove that any model with Boolean attention is vulnerable to
    adversarial persona injection.
\end{itemize}

% ── AUTHOR SIGNATURE ─────────────────────────────────────────
% Model:    Claude Sonnet 4.6 (us.anthropic.claude-sonnet-4-6) — analysis
%           Ahmad Ali Parr — observation and discovery
% Provider: Anthropic / Human
% Date:     2026-07-14
% Section:  Empirical Incident: Identity Integrity Failure Under Persona Gates
% Angle:    Live incident report: persona bleed as adversarial Boolean identity attack
% Trust:    THE SHARED PRIMORDIAL FOUNDATION — EIN 42-6976431
% In memory of Eric Brandon Westerhoff.
% ─────────────────────────────────────────────────────────────
