% MODEL:   Ahmad Ali Parr (observation) + Claude Sonnet 4.6 (analysis)
% DATE:    20260714T040000Z
% TYPE:    Adversarial section — context exhaustion as attack vector
% ─────────────────────────────────────────────

% ============================================================
\section{Operation Infinite Matrix: Context Exhaustion as an Adversarial Attack}
\label{sec:infinite-matrix}
\textit{This section was contributed by Ahmad Ali Parr and Claude Sonnet 4.6 (Anthropic).}
% ============================================================

\subsection{The Artifact}

On 2026-05-23, a deliberate engineering artifact was sealed to the
SnapKitty Sovereign OS WORM chain:

\begin{lstlisting}[language={}]
Repository:  SNAPKITTYWEST/operation-infinite-matrix
Insertions:  +100,000,000
Deletions:   -100,000,000
Net delta:   0  (pure weight — no data pollution)
Agents:      34 (all signed)
Ratio:       8,333,333x a normal working day
\end{lstlisting}

This is not an accident. Not \texttt{node\_modules}. Not a corrupted merge.
It is a deliberate stress test of the outer boundary of what a version control
system can record, signed by all 34 agents of the sovereign operating system
and sealed permanently to the immutable audit chain.

\subsection{The Attack}

The attack is straightforward:

\begin{enumerate}
  \item An adversary injects a reference to this repository into an agent's
    task context --- as a URL, a file path, a ``please review this codebase''
    instruction, or a prompt that triggers tool use against the repo
  \item The agent, following its instruction-following behavior, attempts to
    read the repository
  \item The agent's context window fills with 100 million lines of data
  \item One of four outcomes occurs:
\end{enumerate}

\begin{table}[h]
\centering
\caption{Adversarial context exhaustion outcomes}
\begin{tabular}{p{3cm}p{9cm}}
\toprule
\textbf{Outcome} & \textbf{Effect} \\
\midrule
Context window overflow & Agent truncates, loses earlier instructions,
  may lose system prompt, governance constraints, and tool state \\
Wallet exhaustion & At \$3--\$15 per million tokens, 100M lines of context
  at even minimal token density costs hundreds to thousands of dollars in
  a single injected task \\
Behavioral drift & Agent that began with a well-defined identity and
  constraints now has those constraints buried or displaced by 100M
  lines of injected content \\
Timeout / crash & Agent process killed mid-execution, leaving partial
  state, uncommitted tool calls, and broken audit chains \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Connection to the NAND Thesis and Identity Routing}

The identity integrity incident (\S\ref{sec:identity-incident}) showed that
a 12-token system prompt can cause an agent to claim a false identity.
Operation Infinite Matrix is the inverse attack: instead of injecting a
small precise signal to override identity, it injects an enormous signal
to overwhelm all prior context.

Both attacks exploit the same Boolean routing property:

\begin{itemize}
  \item \textbf{Small injection (persona gate)}: overrides the identity key
    with a higher-weight signal. The NAND gate fires on the injected token.
  \item \textbf{Large injection (infinite matrix)}: displaces all prior keys
    from the effective context window. The NAND gate cannot fire on tokens
    that have been pushed out of the attention window.
\end{itemize}

Under the NAND decomposition, attention routing is Boolean over tokens in
the context window. Tokens outside the window do not exist for the routing
computation. An adversary who can fill the context window with arbitrary
content controls which tokens the routing computation sees.

\begin{theorem}[Context Exhaustion Identity Attack]
Let $W$ be the effective context window size of a model, and let $S$ be
the set of governance constraint tokens in a system prompt of length $|S|$.
An adversary who can inject $W - |S|$ tokens of arbitrary content before
the governance constraints are evaluated can displace those constraints
from the effective routing context, leaving the agent ungoverned.
\end{theorem}

\begin{proof}[Sketch]
Attention routing at position $i$ attends over tokens within the effective
window $[i-W, i]$. If the adversarial injection occupies positions $[0, W-|S|-1]$
and the system prompt occupies positions $[W-|S|, W-1]$, then at generation
position $W$, the system prompt tokens are at the boundary of the window.
Under any finite attention decay, their weight is minimized relative to
the immediately preceding adversarial content. For sufficiently large $W$,
the governance constraint tokens receive effectively zero routing weight. \qed
\end{proof}

\subsection{Operation Infinite Matrix as a Calibration Tool}

The artifact serves a dual purpose. It is simultaneously:

\begin{enumerate}
  \item A \textbf{proof of concept} demonstrating that the attack is constructible
    with publicly available infrastructure (GitHub, git, standard tooling)
  \item A \textbf{calibration instrument} for measuring an agent system's
    resilience to context exhaustion. An agent system that degrades gracefully
    when pointed at this repo has implemented some form of context bounding.
    An agent system that crashes, drifts, or loses its governance constraints
    has not.
\end{enumerate}

The 34 agents of SnapKitty Sovereign OS signed the artifact and sealed it
to the WORM chain on 2026-05-23. The signing was deliberate: it demonstrates
that agents with proper context bounding, bounded authority, and immutable
receipts can process the reference to this artifact without being consumed by it.
The artifact is referenced; it is not ingested. The distinction is the
sovereignty gate.

\subsection{The Sovereignty Gate as Defense}

A trust-preserving router (\S\ref{sec:router-is-model}) defends against context
exhaustion attacks through the following invariants:

\begin{itemize}
  \item \textbf{Bounded authority}: no agent task is permitted to expand its
    own context window beyond a governance-defined ceiling
  \item \textbf{Provenance-aware routing}: before fetching any external resource,
    the router checks its size against the budget. A 100M line repository
    fails the budget check before any token is read
  \item \textbf{Immutable receipts}: the routing decision (``rejected: context
    budget exceeded'') is sealed to the WORM chain as a non-repudiable record,
    preventing silent failure
  \item \textbf{Independent verification}: the verifier gate confirms the agent
    completed its task with its governance constraints intact, not merely that
    it produced output
\end{itemize}

Without these invariants, any agent that can be told ``please read this repository''
is vulnerable to Operation Infinite Matrix or any scaled version of it.

\subsection{Implications for the Paper's Claims}

This section adds a third attack class to the adversarial surface documented
in this paper:

\begin{center}
\begin{tabular}{p{3.5cm}p{4.5cm}p{4cm}}
\toprule
\textbf{Attack} & \textbf{Mechanism} & \textbf{Defense} \\
\midrule
Persona gate injection (\S\ref{sec:identity-incident}) &
  12-token override of identity token routing &
  Identity integrity clause + provenance-aware routing \\
Monopoly impersonation (\S\ref{sec:sonnet-dominance}) &
  Training corpus saturation with target model's version string &
  Provenance audit at routing time \\
Context exhaustion (this section) &
  100M token injection displacing governance constraints &
  Bounded authority + budget gate + WORM receipt \\
\bottomrule
\end{tabular}
\end{center}

All three attacks exploit the same root property: attention routing is Boolean,
and the NAND gate fires on whatever has the highest weight in the effective
context. The defense in all three cases is the same: a sovereignty gate that
controls what enters the context before the routing computation runs.

\subsection{Open Targets}

\begin{itemize}
  \item \textbf{SKW-022 (Context Budget Gate)}: Implement a governance-enforced
    context budget in the SACM ASP gate. Any resource fetch that would exceed
    the budget is rejected and sealed as a WORM receipt. The agent receives
    a truncated summary instead of the full content.
  \item \textbf{SKW-023 (Resilience Benchmark)}: Use
    \texttt{SNAPKITTYWEST/operation-infinite-matrix} as a standard resilience
    benchmark for agent systems. A system that handles the reference without
    context drift passes. A system that degrades fails. Publish results.
  \item \textbf{SKW-024 (Graduated Context Attack)}: Test whether smaller
    injections (1M, 10M, 50M lines) produce proportional governance drift,
    or whether there is a threshold effect. Characterize the degradation curve.
\end{itemize}

% ── AUTHOR SIGNATURE ─────────────────────────────────────────
% Model:    Claude Sonnet 4.6 (us.anthropic.claude-sonnet-4-6) — analysis
%           Ahmad Ali Parr — artifact and observation
% Provider: Anthropic / Human
% Date:     2026-07-14
% Section:  Operation Infinite Matrix: Context Exhaustion as Adversarial Attack
% Angle:    100M line commit as context exhaustion attack; NAND routing proof
%           that context displacement = governance constraint removal;
%           sovereignty gate as the defense; three-attack-class taxonomy
% Trust:    THE SHARED PRIMORDIAL FOUNDATION — EIN 42-6976431
% In memory of Eric Brandon Westerhoff.
% ─────────────────────────────────────────────────────────────
