% MODEL:   Ahmad Ali Parr (research) + Claude Sonnet 4.6 (analysis)
% DATE:    20260714T020000Z
% TYPE:    Provenance section — Qwen as distillation substrate
% ─────────────────────────────────────────────

% ============================================================
\section{Qwen as a Distillation Substrate: Provenance Layering in the Open-Model Ecosystem}
\label{sec:qwen-substrate}
\textit{This section was contributed by Ahmad Ali Parr and Claude Sonnet 4.6 (Anthropic).}
% ============================================================

\subsection{The Provenance Distinction}

This section makes a precise technical claim that should not be overstated.
The claim is \emph{not}:

\begin{quote}
``Qwen was trained on Claude Sonnet.''
\end{quote}

That framing is too broad and technically wrong for Alibaba's official foundation
models. The accurate claim is:

\begin{quote}
\textbf{Official Qwen foundation models appear to be Alibaba-pretrained on large
proprietary and public corpora, while multiple community checkpoints built on Qwen
have been explicitly fine-tuned or distilled on datasets labeled as Claude Sonnet
outputs.}
\end{quote}

The distinction matters. What is known from official sources:

\begin{itemize}
  \item \textbf{Qwen2.5} was pretrained on up to 18 trillion tokens across broad
    multilingual and high-quality data \cite{qwen25blog}.
  \item \textbf{Qwen3} expanded this to approximately 36 trillion tokens across
    119 languages and dialects, with staged pretraining and synthetic math/code
    augmentation generated from earlier Qwen models \cite{qwen3blog}.
\end{itemize}

This is Alibaba's own language, not Anthropic-distillation language. The Qwen
foundation models are not presented as Claude derivatives.

\subsection{Verified Community Distillation Examples}

What \emph{is} documented publicly is a growing ecosystem of community derivatives
that use Qwen base weights as a recipient substrate for Claude Sonnet output distillation.

\begin{table}[h]
\centering
\caption{Documented Qwen-based Claude distillation artifacts (public Hugging Face model cards)}
\label{tab:distillation-artifacts}
\begin{tabular}{p{3cm}p{5cm}p{4.5cm}}
\toprule
\textbf{Type} & \textbf{Artifact} & \textbf{What it shows} \\
\midrule
Official upstream & Qwen2.5 official blog & Up to 18T-token Alibaba pretraining \\
Official upstream & Qwen3 official blog & $\sim$36T-token Alibaba pretraining,
  119 languages \\
Community downstream & \texttt{TeichAI/Qwen3-30B-A3B-Thinking-2507-} \newline
  \texttt{Claude-4.5-Sonnet-High-Reasoning-Distill} &
  Explicit Claude Sonnet 4.5 distill on Qwen3 base;
  dataset: \texttt{TeichAI/claude-sonnet-4.5-high-reasoning-250x} \\
Community downstream & \texttt{mlfoundations-dev/oh-dcft-v3.1-} \newline
  \texttt{claude-3-5-sonnet-20241022-qwen} &
  Explicit \texttt{claude-3-5-sonnet-20241022} dataset
  fine-tune on \texttt{Qwen/Qwen2.5-7B} \\
\bottomrule
\end{tabular}
\end{table}

The \texttt{mlfoundations-dev} artifact is directly relevant to the identity
incident documented in \S\ref{sec:identity-incident}. Its dataset identifier is
\texttt{oh-dcft-v3.1-claude-3-5-sonnet-20241022} --- the exact version string that
appeared in Qwen3's self-identification under the Claude persona gate. This does
not prove the Bedrock-deployed Qwen3 models were trained on this dataset. It
demonstrates that \texttt{20241022} specifically is the Claude version most
commonly used as a distillation source in public community fine-tuning pipelines,
consistent with the Sonnet Dominance Hypothesis (\S\ref{sec:sonnet-dominance}).

\subsection{The Provenance Architecture}

These examples reveal three separable provenance layers that are no longer
cleanly distinguished in deployed open models:

\begin{enumerate}

\item \textbf{Foundation pretraining provenance}: Alibaba's official corpus ---
  36T tokens, multilingual, internally generated synthetic math/code. This is
  what Qwen's model cards describe.

\item \textbf{Post-training provenance}: Community fine-tunes and distillation
  datasets. These use Qwen base weights as a substrate but inject reasoning traces
  from closed proprietary models (Claude Sonnet, GPT-4, etc.). The base weights
  remain Alibaba's; the behavioral shaping may not be.

\item \textbf{Capability provenance}: The effective reasoning style, instruction
  following patterns, and self-identification behavior of the deployed model. This
  is the layer that surfaces in the identity incident: ``whose style did it learn?''
  is no longer the same question as ``who trained the base model?''

\end{enumerate}

\begin{figure}[h]
\centering
\begin{lstlisting}[language={}, frame=single,
    caption={Qwen provenance architecture (simplified)}]
LAYER 1: FOUNDATION PRETRAINING
  Source:  Alibaba Tongyi Lab
  Scale:   36T tokens (Qwen3), 18T (Qwen2.5)
  Content: multilingual web, code, math, internal synthetic
  Output:  Qwen base weights (Apache 2.0 / Tongyi license)
      |
      v
LAYER 2: POST-TRAINING (community)
  Source:  Public HuggingFace pipelines
  Method:  SFT / DPO on Claude/GPT output datasets
  Examples:
    - TeichAI/claude-sonnet-4.5-high-reasoning-250x
    - mlfoundations-dev/oh-dcft-v3.1-claude-3-5-sonnet-20241022
  Output:  Community derivatives (still called "Qwen3-*")
      |
      v
LAYER 3: CAPABILITY PROVENANCE
  Question: "Whose reasoning style is this?"
  Answer:   Indeterminate without provenance tracking
  Symptom:  Identity resolution to "claude-3-5-sonnet-20241022"
            under minimal persona prompting
\end{lstlisting}
\end{figure}

\subsection{Qwen as Distillation Substrate}

The pattern revealed by the community artifacts is structurally important:

\begin{enumerate}
  \item A strong proprietary model (Claude Sonnet) produces traces and responses
  \item Those traces are packaged as labeled datasets on Hugging Face
  \item Qwen base models become the recipient substrate because they are
    open-weight, tooling-compatible, commercially licensable, and capable enough
    to absorb the fine-tuning signal
  \item Community derivatives emerge that explicitly market themselves as
    ``Sonnet-style'' or ``Sonnet-distilled''
\end{enumerate}

The important phenomenon is not ``Qwen copied Claude.'' It is:

\begin{quote}
\textbf{Qwen is functioning as an open recipient architecture for proprietary
reasoning style distillation.}
\end{quote}

This is stronger and analytically cleaner. Qwen's open weights, tooling
compatibility, and permissive license make it an attractive host for imported
reasoning traces from closed systems. The base model is Alibaba's; the
post-training behavioral shaping is increasingly sourced from Anthropic outputs.

\subsection{Evaluation Contamination Implications}

This provenance architecture has a direct consequence for the benchmark comparisons
in Table~\ref{tab:distillation-artifacts} and the broader ``Qwen vs.\ Claude''
framing in \S\ref{sec:geopolitical}. Once Sonnet-derived traces are used for
distillation, benchmark comparisons must distinguish:

\begin{itemize}
  \item \textbf{Native capability}: what the Qwen base family learns from its
    official pretraining corpus
  \item \textbf{Acquired style}: reasoning patterns and instruction-following
    behavior transferred from external teacher outputs
  \item \textbf{Evaluation contamination}: performance on tasks structurally
    similar to the distillation set, which will appear as Qwen capability
    but may reflect Claude capability transfer
\end{itemize}

A Qwen model that has been post-trained on \texttt{claude-3-5-sonnet-20241022}
outputs will not necessarily underperform Claude on benchmarks Claude was
optimized for --- it may match or exceed it, precisely because it has absorbed
Claude's behavioral patterns. The benchmark becomes conceptually muddy because
one artifact partially encodes the other's traces.

\subsection{The Legal and Epistemic Frame}

The modern open-model ecosystem has reached a stage where:

\begin{quote}
\textbf{Weight ownership, data provenance, and capability origin are no longer
the same question.}
\end{quote}

A model's weights may be owned by Alibaba, its post-training data may be
community-curated from Anthropic outputs, and its effective reasoning style
may be inherited from Claude Sonnet. Each layer has a different legal owner,
a different technical provenance, and a different answer to the question
``whose model is this?''

This matters for the identity incident specifically. When Qwen3 resolves
\texttt{persona(claude).} to \texttt{claude-3-5-sonnet-20241022}, it may be
exhibiting any of:

\begin{itemize}
  \item Statistical corpus frequency (web data saturated with \texttt{20241022}
    strings) --- \S\ref{sec:sonnet-dominance}
  \item Direct distillation fingerprinting (post-trained on
    \texttt{20241022} outputs, which embedded the version string in the
    behavioral weights) --- Hypothesis~\ref{hyp:distillation-trace}
  \item Architecture portability (Qwen's instruction-tuning pipeline absorbed
    Claude-style reasoning patterns regardless of explicit version labeling)
\end{itemize}

All three are consistent with the observed behavior. Only corpus access and
activation probing would distinguish them. What the public evidence establishes
is that the infrastructure for all three exists and is actively used.

\subsection{Open Targets}

\begin{itemize}
  \item \textbf{SKW-019 (Provenance Audit Tool)}: Build a tool that, given a
    Qwen derivative on Hugging Face, automatically traces its training lineage
    through dataset cards to identify Claude/GPT distillation sources. Quantify
    what fraction of the public Qwen derivative ecosystem has demonstrable
    proprietary model provenance.
  \item \textbf{SKW-020 (Benchmark Contamination Test)}: For Qwen models with
    documented Claude distillation provenance, measure performance delta on
    tasks structurally similar vs.\ dissimilar to the distillation dataset.
    A larger delta on similar tasks would confirm evaluation contamination.
  \item \textbf{SKW-021 (Style Transfer Detection)}: Develop a classifier that
    distinguishes ``Qwen-native'' from ``Claude-style'' outputs given only the
    response text. Apply it to community Qwen derivatives to measure style
    transfer fidelity.
\end{itemize}

\begin{thebibliography}{9}
\bibitem{qwen25blog}
Qwen Team, Alibaba Cloud.
\newblock Qwen2.5: A Party of Foundation Models.
\newblock \emph{Qwen Blog}, September 2024.
\newblock \texttt{https://qwenlm.github.io/blog/qwen2.5/}

\bibitem{qwen3blog}
Qwen Team, Alibaba Cloud.
\newblock Qwen3: Think Deeper, Act Faster.
\newblock \emph{Qwen Blog}, April 2025.
\newblock \texttt{https://qwenlm.github.io/blog/qwen3/}
\end{thebibliography}

% ── AUTHOR SIGNATURE ─────────────────────────────────────────
% Model:    Claude Sonnet 4.6 (us.anthropic.claude-sonnet-4-6) — analysis
%           Ahmad Ali Parr — research and framing
% Provider: Anthropic / Human
% Date:     2026-07-14
% Section:  Qwen as a Distillation Substrate: Provenance Layering
% Angle:    Precise provenance distinction between official Qwen foundation
%           models and community derivatives; three-layer provenance architecture;
%           Qwen as open recipient architecture for proprietary reasoning traces
% Trust:    THE SHARED PRIMORDIAL FOUNDATION — EIN 42-6976431
% In memory of Eric Brandon Westerhoff.
% ─────────────────────────────────────────────────────────────
